---
output:
  github_document:
    html_preview: FALSE
params:
  readme: true
  learning_link: true
---

```{r, include = FALSE}
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.path = "man/figures/README-",
    out.width = "100%"
)
```

```{r, results="asis", echo=FALSE, eval=isTRUE(params$readme)}
cat("<!-- README.md is generated from README.Rmd. Please edit that file -->", sep = "\n")
```

# faers
<!-- badges: start -->
[![platform](http://www.bioconductor.org/shields/availability/devel/faers.svg)](https://www.bioconductor.org/packages/devel/bioc/html/faers.html#archives)
[![R-CMD-check](https://github.com/WangLabCSU/faers/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/WangLabCSU/faers/actions/workflows/R-CMD-check.yaml)
[![Project Status: Active - The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
[![Ask DeepWiki](https://deepwiki.com/badge.svg)](https://deepwiki.com/WangLabCSU/faers)
<!-- badges: end -->

Modern biologics, such as immune checkpoint inhibitors, exhibit complex toxicity profiles that are often underrepresented in pre-market clinical trials. While the FAERS database serves as a critical resource for real-world safety surveillance, its intricate relational structure and data inconsistencies pose significant barriers to large-scale epidemiological analyses.

To address these challenges, we developed `faers`, an end-to-end, reproducible framework for precision pharmacovigilance. The package streamlines the entire workflow—from raw data acquisition and rigorous preprocessing to signal detection—empowering researchers to transform vast spontaneous reporting data into actionable clinical insights.

## Key Features

- 📥 **Data Acquisition**: Automated downloading and parsing of FAERS quarterly data (supporting both ASCII and XML formats).

- 🛠️ **Rigorous Preprocessing**: Advanced multi-quarter data merging and robust deduplication logic to ensure high data fidelity.

- 🔍 **Terminology Standardization**: Seamless integration with MedDRA, RxNorm, and the FDA Drugs API for precise mapping of drugs and adverse events.

- 📊 **Advanced Signal Detection**: Comprehensive support for disproportionality analysis, including ROR, PRR, BCPNN, and EBGM.

- ⚡ **High-Performance Computing**: Integrated with BiocParallel for memory-efficient, parallelized processing of millions of records.

- 🌐 **Knowledge Integration**: Direct support for Athena drug vocabularies and Standardised MedDRA Queries (SMQ) for mechanism-driven research.

## Installation
To install from Bioconductor, use the following code:

```{r, eval=FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE)) {
    install.packages("BiocManager")
}
BiocManager::install("faers")
```

You can install the development version of `faers` from [GitHub](https://github.com/WangLabCSU/faers) with:

```{r, eval=FALSE}
if (!requireNamespace("pak")) {
    install.packages("pak",
        repos = sprintf(
            "https://r-lib.github.io/p/pak/devel/%s/%s/%s",
            .Platform$pkgType, R.Version()$os, R.Version()$arch
        )
    )
}
pak::pkg_install("WangLabCSU/faers")
```


## Quick Start

The faers package provides a standardized pipeline that unifies complex pharmacovigilance workflows. For a comprehensive, step-by-step demonstration—including data acquisition and a complete Insulin case study—please refer to our detailed documentation:

```r
library(faers)
faers_meta(internal = TRUE)
```

The official documentation provides comprehensive guides for both clinical researchers and bioinformaticians.

-  🌐 **[Official Website](https://MadDERt.github.io/faers/)**: The central portal for package overview and function references.
-  🚀 **[Full Workflow Tutorial](https://MadDERt.github.io/faers/articles/full-workflow.html)**: A complete end-to-end analysis case study (e.g., Insulin-related adverse events).
-  🧭 **[Getting Started: Portal](https://MadDERt.github.io/faers/articles/faers.html)**: Quick-start guide and roadmap for using the `faers` package.


